---
title: 'Chapter 6: Scales, Axes and Legends'
output:
  html_document:
    df_print: paged
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(tidyverse)
library(knitr)

```

# 6.1 Introduction

Scales control the mapping from data to aesthetics. They take data it turn it into something perceivable such as size, color, position, or shape. Scales provide the tools that let a plot to be read: the axes and legends. Formally, each scale is a function from a region in data space (the domain of the scale) to a region in aesthetic space (range of the scale). The axis or legends is the inverse function: it allows for the conversation of visual properties back into data

------------------------------------------------------------------------

# 6.2 Modifying Scales

A scale is required for every aesthetic used on the plot. When writing the following code:

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class))
```

The full code looks like:

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  scale_y_continuous() + 
  scale_x_continuous() + 
  scale_color_discrete()
```

Default scales are named according to the aesthetic and the variable type: scale_y\_continuous(), scale_color_discrete(), etc. ggplot2 automatically adds default scales for every new aesthetic. But to override the defaults, the scale must be manually added

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  scale_x_continuous("The x axis") + 
  scale_y_continuous("Y axis")
```

The use of "+" to "add" scales is misleading. When using "+" to a scale, it's not being added to the plot, but overriding the existing scale

```{r cache = TRUE, fig.show='hold', out.width="50%"}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  scale_x_continuous("Label 1") + 
  scale_x_continuous("Label 2") #  In this case only the second label is applied to the graph

ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  scale_x_continuous("Label 2") # Equivalent plot to the one above
```

A different scale can be used altogether

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  scale_x_sqrt() + 
  scale_color_brewer()
```

The naming schemes for scales is straightforward. It is made up of 3 parts:

1.  scale
2.  The name of the aesthetic (color, shape, or x)
3.  The name of the scale(continuous, discrete, brewer)

------------------------------------------------------------------------

## 6.2.1 Exercises

1.  What happens if you pair a discrete variable to a continuous scale? What happens if you pair a continuous variable to a discrete scale?

```{r}
ggplot(mpg, aes(displ, hwy, color = class)) + 
  geom_point() + 
  scale_x_discrete() + 
  scale_y_discrete() 
```

Scaling a continuous variable on a discrete scale results in the number on the graph being removed and replaced with the name of the variable. Otherwise the graph is similar to as before

```{r eval=FALSE}
ggplot(mpg, aes(drv, hwy)) + 
  geom_point() + 
  scale_x_continuous() + 
  scale_y_continuous()
```

Attempting to scale a discrete variable to a continuous scale will result in an error and the graph will not be produced

2.  Simplify the following plot specifications to make them easier to understand.

```{r eval = FALSE}
ggplot(mpg, aes(displ)) +
  scale_y_continuous("Highway mpg") +
  scale_x_continuous() +
  geom_point(aes(y = hwy))


ggplot(mpg, aes(y = displ, x = class)) +
  scale_y_continuous("Displacement (l)") +
  scale_x_discrete("Car type") +
  scale_x_discrete("Type of car") +
  scale_colour_discrete() +
  geom_point(aes(colour = drv)) +
  scale_colour_discrete("Drive\ntrain")
```

A simplified:

```{r cache=TRUE}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  ylab("Highway mpg")
```

B simplified:

```{r}
ggplot(mpg, aes(class, displ, color = drv)) +
  geom_point() +
  xlab("Type of car") +
  ylab("Displacement (l)") +
  scale_color_discrete("Drive\ntrain")
```

------------------------------------------------------------------------

# 6.3 Guides: Legends and Axes

The most likely component of a scale that is going to modified is **guide**, the axis or legend associated with the scale. Guides allow for observations to be read from the plot and mapped back to their original values. ggplot guides are produced automatically based on the layers of the plot. In ggplot2, the legend isn't directly controlled, instead the data is set up so there's a clear mapping between data and aesthetics that allow for the legend to be generated automatically Axis and legends are the same type of thing but while they look very different there are many natural correspondences between the two:

```{r}
tibble(Axis = c("Label", "Tick and grid line", "Tick label"), Legend = c("Title", "Key", "Key label"), `Argument name` = c("name", "break", "labels"))
```

------------------------------------------------------------------------

## 6.3.1 Scale Title

The first argument to the scale function, *name*, is the axes/legend title. It can be supplied text strings (using \\n for line breaks) or mathematical expressions in quote (as described in ?plotmath)

```{r cache=TRUE, fig.show='hold', out.width="50%"}
df <- data.frame(x = 1:2, y = 1, z = "a")
p <- ggplot(df, aes(x,y)) + geom_point()
p + scale_x_continuous("X axis")
p + scale_x_continuous(quote(a + mathematical ^ expression))
```

Because changing labels is such a common task, there are three helper functions: xlab(), ylab(), and labs():

```{r cache=TRUE, fig.show='hold', out.width="50%"}
p <- ggplot(df, aes(x,y)) + geom_point(aes(color = z))
p + xlab("X axis") + ylab("Y axis") 
p + labs(x = "X axis", y = "Yaxis", color = "Color\nlegend")
```

There are two ways to remove the axis label:

1.  Setting the labs functions to "" omits the label but still allocates space
2.  Setting the labs functions to NULL removes the label and its space

```{r cache=TRUE, fig.show='hold', out.width="50%"}
P <- ggplot(df, aes(x,y)) + geom_point() + theme(plot.background = element_rect(color = "grey50"))
p + labs(x = "", y ="")
p + labs(x = NULL, y = NULL)

```

------------------------------------------------------------------------

## 6.3.2 Breaks and Labels

The **breaks** argument controls which values appear as tick marks on axes and keys on legends. Each break has an associated label, controlled by the **labels** argument. If **labels** is set, than **breaks** must also be set, otherwise the data changes, the breaks will no longer align with the labels:

```{r cache=TRUE, fig.show='hold', out.width="33%"}
df <- data.frame(x = c(1, 3, 5) * 1000, y = 1)
axs <- ggplot(df, aes(x, y)) + geom_point() + labs(x = NULL, y = NULL)
axs
axs + scale_x_continuous(breaks = c(2000, 4000))
axs + scale_x_continuous(breaks = c(2000, 4000), labels = c("2k", "4k"))
```
## 6.3.3 Exercises

1.  Recreate the following graph

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  scale_x_continuous(breaks = c(2,3,4,5,6,7), labels = c("2 L", "3 L", "4 L", "5 L", "6 L", "7 L")) +
  scale_y_continuous(quote(Highway (Miles/Gallon)))
```

2.  List the three different types of object you can supply to the breaks argument. How do breaks and labels differ?

Breaks can be supplied a NULL, a waiver(), a character vector of breaks, or a function. Breaks are the actual values of the plot will scale, while labels are simply the aesthetic display

3.  Recreate the following plot

```{r}
ggplot(mpg, aes(displ, hwy, color = drv)) + 
  geom_point() + 
  scale_color_discrete(labels = c("4wd", "fwd", "rwd"))
```

4.  What label function allows you to create mathematical expressions?What label function converts 1 to 1st, 2 to 2nd, and so on?

quote() is used for mathematical expressions. scales::ordinal_format() to a convert numbers to a list (1st, 2nd, 3rd)

5.  What are the three most important arguments that apply to both axes and legends? What do they do? Compare and contrast their operation for axes vs. legends.

Limits, breaks, and values

------------------------------------------------------------------------

# 6.4 Legends

While the most important parameters are shared between axes and legends, there are some that only apply to legends. Legends are more complicated than axes because:

1.   A legend can display multiple aesthetics (color and shape) from multiple layers, and the symbol displayed in a legends varies based on the geom used in the layer
2.  Axes always appear in the same place. Legends can appear in different places
3.   Legends have considerably more details that can be altered

------------------------------------------------------------------------

## 6.4.1 Layers and Legends

A legend may need to draw symbols from multiple layers. For example if color was mapped to both points and lines, the keys will show both points and lines. The legends varies based on the plot. By default, a layer will only appear if the corresponding aesthetic is mapped to a variable with aes(). A layer can be prevented from appearing in the legend with show.legend: FALSE. However, show.legend : TRUE forces a layer to appear when it otherwise wouldn't. 

```{r cache=TRUE, fig.show='hold', out.width="50%"}

df <- data.frame(x = 1:3, y = 1:3, z = c("a", "b", "c"))
ggplot(df, aes(y,y)) + 
  geom_point(size = 4, color = "grey20") + 
  geom_point(aes(color = z), size = 2)
ggplot(df, aes(y,y)) + 
  geom_point(size = 4, color = "grey20", show.legend = FALSE) + 
  geom_point(aes(color = z), size = 2)
```


Geoms in the legend can be adjusted to display differently to the geom in the plot. By using the override.aes parameter of guide_legend()

```{r cache=TRUE, fig.show='hold', out.width="50%"}
norm <- data.frame(x = rnorm(1000), y = rnorm(1000))
norm$z <- cut(norm$x, 3, labels = c("a","b", "c"))
ggplot(norm, aes(x, y)) + geom_point(aes(color = z, alpha = 0.1))
ggplot(norm, aes(x, y)) + geom_point(aes(color = z, alpha = 0.1)) + guides(color = guide_legend(override.aes = list(alpha = 1)))
```


ggplot2 tries to use the fewest number of legends to accurately convey the aesthetics in the plot. It does this by combining legends where the same variable is mapped to different aesthetics. If both a color and shape are mapped to the same variable then only a single legend is necessary. 

```{r cache=TRUE, fig.show='hold', out.width="33%"}
ggplot(df, aes(x, y)) + geom_point(aes(colour = z)) # Colored points
ggplot(df, aes(x, y)) + geom_point(aes(shape = z)) # Shaped points
ggplot(df, aes(x, y)) + geom_point(aes(shape = z, colour = z)) # Colored shaped points
```

In order for legends to be merged, they must have the same name. Changing the name of one scale requires that all of them be changed. 

***

## 6.4.2 Legend Layout
