---
title: 'Chapter 6: Scales, Axes and Legends'
output:
  html_document:
    df_print: paged
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(tidyverse)
library(knitr)

```

# 6.1 Introduction

Scales control the mapping from data to aesthetics. They take data it turn it into something perceivable such as size, color, position, or shape. Scales provide the tools that let a plot to be read: the axes and legends. Formally, each scale is a function from a region in data space (the domain of the scale) to a region in aesthetic space (range of the scale). The axis or legends is the inverse function: it allows for the conversation of visual properties back into data

------------------------------------------------------------------------

# 6.2 Modifying Scales

A scale is required for every aesthetic used on the plot. When writing the following code:

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class))
```

The full code looks like:

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  scale_y_continuous() + 
  scale_x_continuous() + 
  scale_color_discrete()
```

Default scales are named according to the aesthetic and the variable type: scale_y\_continuous(), scale_color_discrete(), etc. ggplot2 automatically adds default scales for every new aesthetic. But to override the defaults, the scale must be manually added

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  scale_x_continuous("The x axis") + 
  scale_y_continuous("Y axis")
```

The use of "+" to "add" scales is misleading. When using "+" to a scale, it's not being added to the plot, but overriding the existing scale

```{r cache = TRUE, fig.show='hold', out.width="50%"}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  scale_x_continuous("Label 1") + 
  scale_x_continuous("Label 2") #  In this case only the second label is applied to the graph

ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  scale_x_continuous("Label 2") # Equivalent plot to the one above
```

A different scale can be used altogether

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  scale_x_sqrt() + 
  scale_color_brewer()
```

The naming schemes for scales is straightforward. It is made up of 3 parts:

1.  scale
2.  The name of the aesthetic (color, shape, or x)
3.  The name of the scale(continuous, discrete, brewer)

------------------------------------------------------------------------

## 6.2.1 Exercises

1.  What happens if you pair a discrete variable to a continuous scale? What happens if you pair a continuous variable to a discrete scale?

```{r}
ggplot(mpg, aes(displ, hwy, color = class)) + 
  geom_point() + 
  scale_x_discrete() + 
  scale_y_discrete() 
```

Scaling a continuous variable on a discrete scale results in the number on the graph being removed and replaced with the name of the variable. Otherwise the graph is similar to as before

```{r eval=FALSE}
ggplot(mpg, aes(drv, hwy)) + 
  geom_point() + 
  scale_x_continuous() + 
  scale_y_continuous()
```

Attempting to scale a discrete variable to a continuous scale will result in an error and the graph will not be produced

2.  Simplify the following plot specifications to make them easier to understand.

```{r eval = FALSE}
ggplot(mpg, aes(displ)) +
  scale_y_continuous("Highway mpg") +
  scale_x_continuous() +
  geom_point(aes(y = hwy))


ggplot(mpg, aes(y = displ, x = class)) +
  scale_y_continuous("Displacement (l)") +
  scale_x_discrete("Car type") +
  scale_x_discrete("Type of car") +
  scale_colour_discrete() +
  geom_point(aes(colour = drv)) +
  scale_colour_discrete("Drive\ntrain")
```

A simplified:

```{r cache=TRUE}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  ylab("Highway mpg")
```

B simplified:

```{r}
ggplot(mpg, aes(class, displ, color = drv)) +
  geom_point() +
  xlab("Type of car") +
  ylab("Displacement (l)") +
  scale_color_discrete("Drive\ntrain")
```

------------------------------------------------------------------------

# 6.3 Guides: Legends and Axes

The most likely component of a scale that is going to modified is **guide**, the axis or legend associated with the scale. Guides allow for observations to be read from the plot and mapped back to their original values. ggplot guides are produced automatically based on the layers of the plot. In ggplot2, the legend isn't directly controlled, instead the data is set up so there's a clear mapping between data and aesthetics that allow for the legend to be generated automatically Axis and legends are the same type of thing but while they look very different there are many natural correspondences between the two:

```{r}
tibble(Axis = c("Label", "Tick and grid line", "Tick label"), Legend = c("Title", "Key", "Key label"), `Argument name` = c("name", "break", "labels"))
```

------------------------------------------------------------------------

## 6.3.1 Scale Title

The first argument to the scale function, *name*, is the axes/legend title. It can be supplied text strings (using \\n for line breaks) or mathematical expressions in quote (as described in ?plotmath)

```{r cache=TRUE, fig.show='hold', out.width="50%"}
df <- data.frame(x = 1:2, y = 1, z = "a")
p <- ggplot(df, aes(x,y)) + geom_point()
p + scale_x_continuous("X axis")
p + scale_x_continuous(quote(a + mathematical ^ expression))
```

Because changing labels is such a common task, there are three helper functions: xlab(), ylab(), and labs():

```{r cache=TRUE, fig.show='hold', out.width="50%"}
p <- ggplot(df, aes(x,y)) + geom_point(aes(color = z))
p + xlab("X axis") + ylab("Y axis") 
p + labs(x = "X axis", y = "Yaxis", color = "Color\nlegend")
```

There are two ways to remove the axis label:

1.  Setting the labs functions to "" omits the label but still allocates space
2.  Setting the labs functions to NULL removes the label and its space

```{r cache=TRUE, fig.show='hold', out.width="50%"}
P <- ggplot(df, aes(x,y)) + geom_point() + theme(plot.background = element_rect(color = "grey50"))
p + labs(x = "", y ="")
p + labs(x = NULL, y = NULL)

```

------------------------------------------------------------------------

## 6.3.2 Breaks and Labels

The **breaks** argument controls which values appear as tick marks on axes and keys on legends. Each break has an associated label, controlled by the **labels** argument. If **labels** is set, than **breaks** must also be set, otherwise the data changes, the breaks will no longer align with the labels:

```{r cache=TRUE, fig.show='hold', out.width="33%"}
df <- data.frame(x = c(1, 3, 5) * 1000, y = 1)
axs <- ggplot(df, aes(x, y)) + geom_point() + labs(x = NULL, y = NULL)
axs
axs + scale_x_continuous(breaks = c(2000, 4000))
axs + scale_x_continuous(breaks = c(2000, 4000), labels = c("2k", "4k"))
```
## 6.3.3 Exercises

1.  Recreate the following graph

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  scale_x_continuous(breaks = c(2,3,4,5,6,7), labels = c("2 L", "3 L", "4 L", "5 L", "6 L", "7 L")) +
  scale_y_continuous(quote(Highway (Miles/Gallon)))
```

2.  List the three different types of object you can supply to the breaks argument. How do breaks and labels differ?

Breaks can be supplied a NULL, a waiver(), a character vector of breaks, or a function. Breaks are the actual values of the plot will scale, while labels are simply the aesthetic display

3.  Recreate the following plot

```{r}
ggplot(mpg, aes(displ, hwy, color = drv)) + 
  geom_point() + 
  scale_color_discrete(labels = c("4wd", "fwd", "rwd"))
```

4.  What label function allows you to create mathematical expressions?What label function converts 1 to 1st, 2 to 2nd, and so on?

quote() is used for mathematical expressions. scales::ordinal_format() to a convert numbers to a list (1st, 2nd, 3rd)

5.  What are the three most important arguments that apply to both axes and legends? What do they do? Compare and contrast their operation for axes vs. legends.

Limits, breaks, and values

------------------------------------------------------------------------

# 6.4 Legends

While the most important parameters are shared between axes and legends, there are some that only apply to legends. Legends are more complicated than axes because:

1.   A legend can display multiple aesthetics (color and shape) from multiple layers, and the symbol displayed in a legends varies based on the geom used in the layer
2.  Axes always appear in the same place. Legends can appear in different places
3.   Legends have considerably more details that can be altered

------------------------------------------------------------------------

## 6.4.1 Layers and Legends

A legend may need to draw symbols from multiple layers. For example if color was mapped to both points and lines, the keys will show both points and lines. The legends varies based on the plot. By default, a layer will only appear if the corresponding aesthetic is mapped to a variable with aes(). A layer can be prevented from appearing in the legend with show.legend: FALSE. However, show.legend : TRUE forces a layer to appear when it otherwise wouldn't. 

```{r cache=TRUE, fig.show='hold', out.width="50%"}

df <- data.frame(x = 1:3, y = 1:3, z = c("a", "b", "c"))
ggplot(df, aes(y,y)) + 
  geom_point(size = 4, color = "grey20") + 
  geom_point(aes(color = z), size = 2)
ggplot(df, aes(y,y)) + 
  geom_point(size = 4, color = "grey20", show.legend = FALSE) + 
  geom_point(aes(color = z), size = 2)
```


Geoms in the legend can be adjusted to display differently to the geom in the plot. By using the override.aes parameter of guide_legend()

```{r cache=TRUE, fig.show='hold', out.width="50%"}
norm <- data.frame(x = rnorm(1000), y = rnorm(1000))
norm$z <- cut(norm$x, 3, labels = c("a","b", "c"))
ggplot(norm, aes(x, y)) + geom_point(aes(color = z, alpha = 0.1))
ggplot(norm, aes(x, y)) + geom_point(aes(color = z, alpha = 0.1)) + guides(color = guide_legend(override.aes = list(alpha = 1)))
```


ggplot2 tries to use the fewest number of legends to accurately convey the aesthetics in the plot. It does this by combining legends where the same variable is mapped to different aesthetics. If both a color and shape are mapped to the same variable then only a single legend is necessary. 

```{r cache=TRUE, fig.show='hold', out.width="33%"}
ggplot(df, aes(x, y)) + geom_point(aes(colour = z)) # Colored points
ggplot(df, aes(x, y)) + geom_point(aes(shape = z)) # Shaped points
ggplot(df, aes(x, y)) + geom_point(aes(shape = z, colour = z)) # Colored shaped points
```

In order for legends to be merged, they must have the same name. Changing the name of one scale requires that all of them be changed. 

***

## 6.4.2 Legend Layout

A number of settings that affect the overall display of the legends are controlled through the theme system. Modify the legend with the theme() function. The position and justification of legends are controlled by the theme setting legend.position. It takes the values "right", "left", "top" "bottom", or "none" (no legend)

```{r cache=TRUE, fig.show='hold', out.width="33%"}
df <- data.frame(x = 1:3, y = 1:3, z = c("a", "b", "c")) 
base <- ggplot(df, aes(x, y)) + geom_point(aes(color = z), size = 3) + xlab(NULL) + ylab(NULL)

base + theme(legend.position = "right") # The default
base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "none")
```

Switching between left/right and top/bottom modifies how the keys in each legend are laid out (horizontal or vertically), and how multiple legends are stacked (horizontally or vertically). These options can be adjusted independently:

-   legend.direction - Changes the layout of items in legends ("horizontal" or "vertical")

-   legend.box - Changes arrangement of multiple legends ("horizontal" or "vertical")

-   legend.box - Justification of each legend within the overall bounding box, when there are multiple legends ("top", "bottom", "left", or "right")

Alternatively, if there's a lot of blank space on the plot, the legend can be placed inside the plot. This can be done by setting legend.position to a numeric vector of length two. The number represents a relative location in the panel area: c(0,1) is top left, c(1,0) is bottom-right. The corner of which legend.position refers to can be controlled with legend.justification. However, positioning the legend exactly in a specific spot will require a lot of trial and error

```{r cache=TRUE, fig.show='hold', out.width="25%"}
base <- ggplot(df, aes(x,y)) + geom_point(aes(color = z), size = 3)

base + theme(legend.position = c(0, 1), legend.justification = c(0, 1))

base + theme(legend.position = c(0.5, 0.5), legend.justification = c(0.5, 0.5))

base + theme(legend.position = c(1, 0), legend.justification = c(1, 0))

base + theme(legend.position = c(0.3, 0.7), legend.justification = c(1, 0.6))
```

There is also a margin around the legends that can be suppressed with legend.margin = unit(0, "mm")

------------------------------------------------------------------------

## 6.4.3 Guide Functions

The guide functions, guide_colorbar() and guide_legend(), offer additional control over the fine details of the legend. Legends guides can be used for any aesthetic (discrete or continuous), while the color bar can only be used with continuous color scales. The default guide can be overrided using the guide argument of the corresponding scale function, or more conveniently, the guides() helper function. guides() works like labs(): it can override the default guide associated with each aesthetic

```{r cache=TRUE, fig.show='hold', out.width="33%"}
df <- data.frame(x = 1, y = 1:3, z = 1:3)
base <- ggplot(df, aes(x, y)) + geom_raster(aes(fill = z))
base
base + scale_fill_continuous(guide = guide_legend())
base + guides(fill = guide_legend())

```

Both functions can take numerous arguments to help control the finer level of details of the text, color, size, and font.

------------------------------------------------------------------------

### 6.4.3.1 guide_legend()

The legend guide displays individual keys in a table. The most useful options are:

-   *nrow* or *ncol* which specify the dimensions of the table. Byrow controls how the table is filled: False fills it by column (the default), True fills it by row

```{r cache=TRUE, fig.show='hold', out.width="33%"}
df <- data.frame(x = 1, y = 1:4, z = letters[1:4])
p <- ggplot(df, aes(x,y)) + geom_raster(aes(fill = z))
p
p + guides(fill = guide_legend(ncol = 2))
p + guides(fill = guide_legend(ncol = 2, byrow = TRUE))

```

-   *reverse* reverses the order of the keys. However the default stack is in the correct order for the graph and legend

```{r cache=TRUE, fig.show='hold', out.width="33%"}
p <- ggplot(df, aes(1, y)) + geom_bar(stat = "identity", aes(fill = z))
p
p + guides(fill = guide_legend(reverse = TRUE))
```

-   *override.aes* overrides some of the aesthetic setting derived from each layer. This is useful to make elements in the legend more visually prominent

-   *keywidth* and *keyheight* (along with default.unit) allows for the specification of the size of the keys. These are grid units (Ex. unit(1, "cm"))

------------------------------------------------------------------------

### 6.4.3.2 guide_colorbar

The color bar guide is designed for continuous range of color. It outputs a rectangle over which the color gradient varies. The most important arguments are:

-   *barwidth* and *barheight* (along with *default.unit*) allowing for the specification of the size of the bar. These are grid units (Ex unit(1, "cm"))

-   *nbin* controls the number of slices. Increase this from the default value of 20 if drawing a very long bar

-   *reverse* flips the color bar to the lowest values at the top

```{r cache=TRUE, fig.show='hold', out.width="33%"}
df <- data.frame(x = 1, y = 1:4, z = 4:1)
p <- ggplot(df, aes(x, y)) + geom_tile(aes(fill = z))

p
p + guides(fill = guide_colorbar(reverse = TRUE))
p + guides(fill = guide_colorbar(barheight = unit(4, "cm")))
```

------------------------------------------------------------------------

## 6.4.4 Exercises

1.  How do you make legends appear to the left of the plot?

```{r cache=TRUE, fig.show='hold', out.width="50%"}
ggplot(mpg, aes(displ,hwy, color = drv)) + 
  geom_point() # Original plot
ggplot(mpg, aes(displ,hwy, color = drv)) + 
  geom_point() + 
  theme(legend.position = "left") # Plot on the left side
```

2.  What's gone wrong with this plot? How could you fix it?

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point(aes(colour = drv, shape = drv)) + scale_colour_discrete("Drive train") 


```

There are two separate legends when they should only be one

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point(aes(colour = drv, shape = drv)) + scale_color_discrete("Drive train") + scale_shape_discrete("Drive train")
```

Fixed graph by manually setting both the shape and color to the same title

3.  Can you recreate the code for this plot?

```{r}
ggplot(mpg, aes(displ, hwy, color = class)) + 
  geom_point(size = 3, show.legend = FALSE) + 
  geom_smooth(method = "lm", aes(displ, hwy, color = class), se = FALSE) + 
  theme(legend.position = "bottom", legend.title = element_blank()) +
  guides(color = guide_legend(ncol = 8, byrow = TRUE))
```

------------------------------------------------------------------------

# 6.5 Limits

The limits, or domain, of a scale are usually derived from the range of the data. There are two reasons why it might be better to specify the limits rather than rely on the data:

1.  Making the limits smaller than the range of the data allows for focus on an interesting area of the plot
2.  Making the limits larger than the range of data to allow for multiple plots to match up

Limits are naturally though of as the limits of a position scale, but limits also apply to legends, colors, sizes, and shapes

Limits can be modified with the limits parameter of the scale:

-   For continuous scales, a numeric vector of length two is applied. To set only an upper or lower limit, set the other value to NA

-   For discrete scales, a character vector which enumerates all possible values

```{r cache=TRUE, fig.show='hold', out.width="33%"}
df <- data.frame(x = 1:3, y = 1:3)
base <- ggplot(df, aes(x, y)) + geom_point()


base
base + scale_x_continuous(limits = c(1.5, 2.5)) # Removed 2 rows containing missing values (`geom_point()`). 
base + scale_x_continuous(limits = c(0, 4))
```

Modifying limits is a common tasks, ggplot2 provides helpers such as xlim(), ylim(), and lims(). These functions inspect their input and create an appropriate scale:

-   xlim(10.20) - A continuous scale from 10 to 20

-   ylim(20,10) - A reversed continuous scale from 20 to 10

-   xlim("a", "b", "c") - A discrete scale.

-   xlim(as.Date(c("2008-05-01", "2008-08-01"))): a date scale from May 1 to August 1 2008.

```{r cache=TRUE, fig.show='hold', out.width="33%"}
base + xlim(0, 4)
base + xlim(4, 0)
base + lims(x = c(0, 4))
```

The range of the axes always extend a little bit past the limits that have been specified to ensure data does not overlap with the axes. This can be eliminated with set *expand = c(0,0)*

```{r cache=TRUE, fig.show='hold', out.width="50%"}
ggplot(faithfuld, aes(waiting, eruptions)) + 
  geom_raster(aes(fill = density)) + 
  theme(legend.position = "none")
ggplot(faithfuld, aes(waiting, eruptions)) +
  geom_raster(aes(fill = density)) + scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(legend.position = "none")
```

By default any data outside the limits is converted to NA. This means that setting the limits is not the same as visually zooming into a region of the plot. To zoom zoom into the plot, use xlim and ylim arguments to coord_Cartesian(), this performs purely visual zooming and does not affect the underlying data. This value can also be overrided with oob (out of bounds) argument to the scale. The default is scales::censor() which replaces any value outside the limit with NA. Another option is scales::squish() which squishes all values into the range

```{r cache=TRUE, fig.show='hold', out.width="33%"}
df <- data.frame(x = 1:5)
p <- ggplot(df, aes(x, 1)) + geom_tile(aes(fill = x), colour = "white")
p
p + scale_fill_gradient(limits = c(2, 4))
p + scale_fill_gradient(limits = c(2, 4), oob = scales::squish)

```

------------------------------------------------------------------------

## 6.5.1 Exercises

1.  The following code creates two plots of the mpg dataset. Modify the code so that the legend and axes match, without using faceting

```{r cache=TRUE, fig.show='hold', out.width="50%"}
fwd <- subset(mpg, drv == "f")
rwd <- subset(mpg, drv == "r")

ggplot(fwd, aes(displ, hwy, colour = class)) + geom_point()
ggplot(rwd, aes(displ, hwy, colour = class)) + geom_point()
```

Solution:

```{r cache=TRUE, fig.show='hold', out.width="50%"}

fwd <- subset(mpg, drv == "f")
rwd <- subset(mpg, drv == "r")

ggplot(fwd, aes(displ, hwy, colour = class)) + 
  geom_point() +
  scale_color_discrete("Drive train") +
  xlim(0, 10) +
  ylim(0, 45) +
  expand_limits(color = c("2seater", "compact", "midsize", "minivan", "pickup", "subcompact", "suv"))
ggplot(rwd, aes(displ, hwy, colour = class)) + 

    geom_point() +
  scale_color_discrete("Drive train")+
  xlim(0, 10) +
  ylim(0, 45) +
  expand_limits(color = c("2seater", "compact", "midsize", "minivan", "pickup", "subcompact", "suv"))
```

2.  What does expand limits() do and how does it work?

The function takes a values or values and expands to the graph to include a singular point or a given range.

```{r cache=TRUE, fig.show='hold', out.width="50%"}
p <- ggplot(mtcars, aes(mpg, wt)) + geom_point()
p
p + expand_limits(x = 0) # The point zero is included in the graph
p + expand_limits(y = c(1, 15)) # includes all y values between 1 and 15
p + expand_limits(x = 0, y = 0) # Includes the origin in the set of data
```

3.  What happens if you add two xlim() calls to the same plot? Why?

```{r}
p + xlim(15,30) + xlim(10, 40) # The existing scale is replaced with the scale that comes after it

```

4.  What does scale x continuous(limits = c(NA, NA)) do?

```{r}
p <- ggplot(mtcars, aes(mpg, wt)) + geom_point()
p + scale_x_continuous(limits = c(NA, NA))
```

It simply uses the original limits assigned from the data

------------------------------------------------------------------------

# 6.6 Scales Toolbox

Besides adjusting the options of the default scales, they can also be completely replaced with new scales. Scales can be divided roughly into four families:

-   Continuous position scales used to map integer, numeric, and date/time data to x and y positions

-   Color scales, used to map continuous and discrete data to colors

-   Manual scales used to map discrete variables to an aesthetic such as size, line, type, shape, or color.

-   The identity scale, paradoxically used to plot variables without scaling them.

------------------------------------------------------------------------
## 6.6.1 Continuous Position Scales

Every plot has two position scales, *x* and *y*. The most common continuous position scales are scale_x\_continuous() and scale_y\_continuous() which linearly map data to the x and y axis. Interesting variations of these scales are produced through transformations. Every continuous scale takes a *trans* argument, allowing for a variety of transformations:

```{r cache=TRUE, fig.show='hold', out.width="25%"}
ggplot(mpg, aes(displ, hwy)) + geom_point() 
ggplot(mpg, aes(displ, hwy)) + geom_point() + scale_y_continuous(trans = "reciprocal") # Converts fuel economy to fuel consumption

ggplot(diamonds, aes(price,carat)) + geom_bin2d()
ggplot(diamonds, aes(price, carat)) + geom_bin2d() + scale_x_continuous(trans = "log10") +scale_y_continuous(trans = "log10") # Log10 x and y axis
```

The transformation is carried out by a *transformer*, which describes the transformation, its inverse, and how to draw the labels. Common shortcuts for transformers include: scale_x\_log10(), scale_x\_sqrt(), and scale_x\_reverse()

Transformation can also be done on the data instead of scaling the plot. If using using a transformed scale, the axes will be labelled in the original data space, if the data is transformed the axes will be labelled in the transformed space. Transformation always occurs before statistical summaries. To transform, after statistical computations use *coord_trans()*

Date and date/time data are continuous variables with special labels. scale_x\_date() and scale_x\_datetime() work similarly to scale_x\_continuous() but have special arguments that work in date_friendly units: - *date_breaks* and *date_minor_breaks()* allows for the positioning by date units (years, months, weeks, day, hours, minutes, and seconds) - date_labels controls the display of the labels using the same formatting strings as in strptime() and format()

```{r cache=TRUE, fig.show='hold', out.width="25%"}
base <- ggplot(economics, aes(date, psavert)) + geom_line(na.rm = TRUE) + labs(x = NULL, y = NULL)
base # Default breaks and labels
base + scale_x_date(date_labels = "%y", date_breaks = "5 years")

base + scale_x_date(
  limits = as.Date(c("2004-01-01", "2005-01-01")),
  date_labels = "%b %y",
  date_minor_breaks = "1 month"
)
base + scale_x_date(
  limits = as.Date(c("2004-01-01", "2004-06-01")),
  date_labels = "%m/%d",
  date_minor_breaks = "2 weeks"
)
  
```

------------------------------------------------------------------------

## 6.6.2 Color

############################################################################################################################################################################################### 

After position, the most commonly used aesthetic is color. HCL (Hue, chroma, and luminescence) color space has three components:

1.   Hue, a number between 0 and 360 (an angle) which gives the color of the color
2.  Chroma is the purity of a color. A chroma of 0 is grey, and the maximum value of chroma varies with luminescence.
3.  Luminescence is the lightness of the color. A luminescence of 0 produces black and a luminescence of 1 produces white

